---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Reproducible research Week 2: Assignment 
=========================================

```{r, echo=FALSE, results='hide'}
time <- format(Sys.time(), "%Y-%m-%d")
```

Today is `r time`.

Clear the workspace and load packages. 
```{r, results='hide', warning=FALSE, comment=FALSE, message=FALSE}
rm(list=ls())
library(dplyr)
library(tidyr)
```


Download and upload data. 

```{r, results='hide'}
wd <- "C:/Users/vi04wecu/Dropbox/Projects/ML/R/Reproducible_research/Week_2/RR_week_2"
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
myfile<- "data.zip"   

setwd(wd)

download.file(url, myfile, mode="wb")
unzip(myfile)
list.files(wd)
data<- read.csv("activity.csv")

```


Data manipulation and description

```{r}
data$date<- as.Date(data$date, "%Y-%m-%d")
data$steps<- as.numeric(data$steps)
data$interval<- as.numeric(data$interval)
str(data)
summary(data)
datanotmissing <- data[!is.na(data$steps), ]
```

##What is mean total number of steps taken per day?
Make a histogram of the total number of steps taken each day

```{r}

sumsteps<-aggregate(datanotmissing$steps, by=list(Category=datanotmissing$date), FUN=sum)
hist(sumsteps$x, xlab = "Total number of steps per day", main="Total Number of Steps per Day")
```

Calculate and report the mean and median total number of steps taken per day

```{r}
mean<- mean(sumsteps$x,  na.rm=TRUE)
median <- median(sumsteps$x,  na.rm=TRUE)
```

Mean of total number of steps is **`r mean`**, median is **`r median`**.

##What is the average daily activity pattern? 

```{r}
meansteps<-aggregate(datanotmissing$steps, by=list(Category=datanotmissing$interval), FUN=mean)

```

```{r}
plot(x=meansteps$Category, y=meansteps$x, type = "l", main = "Average daily activity pattern", ylab = "Number of Steps", xlab = "5-minute interval")
```
 
 
```{r}
max <- max(meansteps$x)
maxint<- meansteps[meansteps$x == max , 1]
```
 
The maximum number of steps is in the **`r maxint`** interval.

##Imputing missing values

```{r}
summissing<- sum(is.na(data))
```
There are **`r summissing`** missing values in the dataset.


```{r}
dataimputed <- data

dataimputed<- dataimputed %>% group_by(interval) %>% mutate(num_avg = mean(steps,na.rm=T))
dataimputed$steps[is.na(dataimputed$steps)] <- dataimputed$num_avg[is.na(dataimputed$steps)]
```


```{r}
sumsteps_imp<-aggregate(dataimputed$steps, by=list(Category=dataimputed$date), FUN=sum)
hist(sumsteps_imp$x, xlab = "Total number of steps per day", main="Total Number of Steps per Day")
```

```{r}
mean_imp<- mean(sumsteps_imp$x,  na.rm=TRUE)
median_imp <- median(sumsteps_imp$x,  na.rm=TRUE)
```

Mean of total number of steps is **`r mean`** , median is **`r median`**.

The calculations are the same.

##Are there differences in activity patterns between weekdays and weekends?

```{r}
dataimputed$weekdays<- weekdays(dataimputed$date)

 dataimputed$weekend<- as.numeric(rep(0, times=length(dataimputed$steps)))
 for (x in 1:length(dataimputed$steps)) {
   
   if (dataimputed[x, 5] %in% c("Samstag", "Sonntag")) {
     dataimputed[x, 6]<- 1
   }
   
 }
 
 weekdaydata<- dataimputed[dataimputed$weekend==0, ]
 meansteps_weekday<-aggregate(weekdaydata$steps, by=list(Category=weekdaydata$interval), FUN=mean)

 
 weekendata<- dataimputed[dataimputed$weekend==1, ]
 meansteps_weekend<-aggregate(weekendata$steps, by=list(Category=weekendata$interval), FUN=mean)
```


Plot by group

```{r}
par(mfrow=c(2,1),  mar=c(4, 4, 3, 2))
plot(x=meansteps_weekday$Category, y=meansteps_weekday$x, type = "l", main = "Average daily activity pattern (Weekday)", ylab = "Number of Steps", xlab = "5-minute interval")

plot(x=meansteps_weekend$Category, y=meansteps_weekend$x, type = "l", main = "Average daily activity pattern (Weekend)", ylab = "Number of Steps", xlab = "5-minute interval")


dev.off()  
```

